---
title: "Task 1 - Spatial Data Visualization"
author: "Matthieu Huy"
date: "2023-03-14"
output: 
  html_document:
    code_folding: hide
---

```{r setup, echo = TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(broom)
library(janitor)

# Spatial data packages
library(sf)
library(tmap)
```

For this analysis we will use two datasets:\

- California Department of Fish and Wildlife. (2020) Oil Spill Incident Tracking, California, 2008 (oil_spill_incident_tracking_[ds394].csv)(accessed [here](https://gis.data.ca.gov/datasets/7464e3d6f4924b50ad06e5a553d71086_0/explore?location=36.780007%2C-119.422009%2C6.74)).\

- California County shapefile (ca_counties_tiger_2016.shp) from the US Census Bureau's 2016 MAF/TIGER database (accessed [here](https://data.ca.gov/dataset/ca-geographic-boundaries/resource/b0007416-a325-4777-9295-368ea6b710e6?inner_span=True).)\


The DFW covers all of the state of California. Each point in this dataset represents an individual oil spill incident defined as, "a discharge or threatened discharge of petroleum or other deleterious material into the waters of the state."\


```{r}

#### Download, read, and transform data

### read in Oil Spill Incidents csv:

oil_spills_df <- read_csv(here('data/oil_spill_incident_tracking_[ds394].csv')) %>% 
  clean_names()

#remove 00:00:00+00 from date for tmap labelling purposes
oil_spills_df$dateofinci <- str_remove(oil_spills_df$dateofinci, "00:.+") 

### read in CA county shapefile:
ca_counties_sf <- read_sf(here('data/ca_counties/CA_Counties_TIGER2016.shp')) %>% 
  clean_names() %>% 
  select(namelsad, countyfp) #reorder so that county name is first column (for tmap)


### convert oil_spills_df to sf:
oil_spills_sf <- st_as_sf(oil_spills_df,  ### convert to sf
                          coords = c('x', 'y'), ### specify spatial information for geometry
                          crs = st_crs(ca_counties_sf)) ### set coord system same as county_sf
```

```{r plot1, results = 'asis', out.width = "100%", fig.align = 'center', fig.cap = "**Figure 1: Oil spill incidents throughout California counties in 2008.** Points represent individual oil spill events."}

tmap_mode(mode = "view") ### set mode to view/interactive

plot1 <- tm_shape(ca_counties_sf) + ### county polygons
      tm_borders("lightgrey", lwd = 0.5) + ###border color and width
  tm_shape(oil_spills_sf) + ### oil spill points
  tm_dots(id = "dateofinci", ### label with dates instead of point id number
          col = "darkcyan", ### color points
          size = 0.005) ### size of points

tmap_leaflet(plot1) ### render plot1 tmap
```
\
\
\
```{r, fig.align = 'center', out.width = "100%"}

### perform spatial join of oil_spills_sf to ca_counties_sf
county_oil_spills_sf <- ca_counties_sf %>% 
  st_join(oil_spills_sf) %>% 
  filter(inlandmari %in% c("Inland")) #keep only inland oil spills

###check for NA values
#sum(is.na(county_oil_spills_sf$oesnumber)) 
###no NA values

###count number of oil spill observations for each county
oil_spill_counts_sf <- county_oil_spills_sf %>% 
  count(namelsad)

legend_breaks <- c("25","50","75","100","125","150","175","200+")

ggplot() +
  geom_sf(data = ca_counties_sf, #to fill missing county in top right corner of map
          aes(color = "lightgrey"),
          color = "white",
          size = 0.1) +
  geom_sf(data = oil_spill_counts_sf,
          aes(fill = n,),
          color = "white",
          size = 0.1) +
  scale_fill_stepsn(limits = c(0, 220), breaks = c(20, 40, 60, 80, 100, 120, 140, 160, 180, 200), 
                    labels = c("20", "40", "60", "80", "100", "120", "140", "160", "180", "200+"),
                    colors = hcl.colors(8, palette = "Inferno", rev = TRUE)) +
  theme_void() +
  labs(fill = "",
       title = "Figure 2: Number of oil spill events in California counties in 2008") +
  theme(plot.title = element_text(size = 10, color = "black", face = "bold"),
        legend.title = element_text(size = 10))
```

